An FMI-Based Initialization Plugin for INTO-CPS Maestro 2

### 2020-06-29 background.tex
Added some based on the feedback from Christian

\begin{definition}[Algebraic loops] 
The set of interconnected variables in an algebraic loop is defined as: $LoopVariables = \{v | v \in V : Path(v,v)\}$ \\
An algebraic loop is defined as the set $\{p_1, p_2 | p_1, p_2 \in LoopVariables: Path(p_1, p_2) \land Path(p_1, p_2)\}$\todo{Look at this}\\
Path is defined as the transitive closure of the set of relations: $\bigcup_{c \in \fmus} \set{\feedthrough{c}} \cup \coupling$\\
The algebraic dependency will lead to the following difference between performing to following :
\todo{Claudio check this definition and add a formalization of the challenge introduced}
\end{definition}

@@

As described by definition \ref{def:challenge} the algebraic loops introduce certain challenges in the system. This means that algebraic loops needs to be handled using fixed point iterations\cite{Gomes2018}. It is a technique to repeatedly perform the steps of a sub-list of the co-simulation step. The number of repetitions the operations should be be performed depends on the characteristics of the scenario. The operations should be performed until the system converge.
However a fixed point iteration technique is not guaranteed to convergence if the system is unstable. This means that an upper bound of the number of repetitions needs to be established to ensure termination. In case of a non-converging algebraic loop this should be treated as an incorrect co-simulation scenario and the simulation should be stopped. The criteria of a valid co-simulation scenario is is specified in definition \ref{def:convergence}.

\todo{Christian should I add some pseudo code to explain?}
\begin{definition}[Convergence of Fixed point iteration]\label{def:convergence}
A fixed point iteration is converging if a finite number of iterations will make the difference of the output-value of the same operation between two following iterations within a certain threshold.\\
$Convergence = \exists n \in \setnat: \forall p \in LoopVariables \implies |f(p, n + 1) - f(p, n)| \leq T$\\
$f$ is the operation performed on a loop-variable p of a given iteration n.
\end{definition}

>>>>

\begin{definition}[Algebraic loops] 
The set of interconnected variables in an algebraic loop is defined as: $LoopVariables = \{v | v \in V : Path(v,v)\}$ \\
An algebraic loop is defined as the set $\{p_1, p_2 | p_1, p_2 \in LoopVariables: Path(p_1, p_2) \land Path(p_2, p_1)\}$\todo{Look at this - can it be specified better?}\\
Path is defined as the transitive closure of the set of relations: $\bigcup_{c \in \fmus} \set{\feedthrough{c}} \cup \coupling$\\
The algebraic dependency will lead to the following difference between performing to following :
\todo{Claudio check this definition and add a formalization of the challenge introduced}
\end{definition}

@@

As described by definition \ref{def:challenge} the algebraic loops introduce certain challenges in the system. This means that algebraic loops needs to be handled using fixed point iterations\cite{Gomes2018}. It is a technique to repeatedly perform the steps of a sub-list of the co-simulation step. The number of repetitions the operations should be be performed depends on the characteristics of the scenario. The operations should be performed until the system converge. 
An example of applying a fixed point iteration can be seen in Figure \ref{} where a bigger system contains an algebraic loop.

\begin{figure}[H]
    \centering
    \includegraphics{images/fmu_cycle.pdf}
    \caption{Co-simulation scenario with cycle}
    \label{fig:fmu_cycle}
\end{figure}

The initialization of the system requires application of fixed point iteration as defined by definition \ref{def:fixedpoint}.

\begin{definition}[Fixed point iteration of an Initialization procedure]\label{def:fixedpoint}
Given an Initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ of a co-simulation $\tuple{\fmus, \coupling}$ where $LoopVariables \neq \emptyset$.\\
Fixed point iteration should be used by 
$\forall v \in LoopVariables $

\end{definition}

However a fixed point iteration technique is not guaranteed to convergence if the system is unstable. This means that an upper bound of the number of repetitions needs to be established to ensure termination. In case of a non-converging algebraic loop the simulation should be stopped since the result of the co-simulation scenario wouldn't be trustworthy. The criteria of a valid co-simulation scenario is is specified in definition \ref{def:convergence}.

\begin{definition}[Convergence of Fixed point iteration]\label{def:convergence}
A fixed point iteration is converging if a finite number of iterations will make the difference of the output-value of the same operation between two following iterations within a certain threshold.\\
$Convergence = \exists n \in \setnat: \forall p \in LoopVariables \implies |f(p, n + 1) - f(p, n)| \leq T$\\
$f$ is the operation performed on a loop-variable p of a given iteration n.
\end{definition}

### 2020-06-29 initializationorder.tex
Added some based on the feedback from Christian

\section{Calculation of an Initialization Order}\label{sc:initilization}
FMI defines certain information about the initialization order described through different states of a co-simulation. The initialization phase covers the two states (in chronological order):
\begin{itemize}
    \item \textit{Instantiated}
    \item \textit{Initialization Mode}
\end{itemize}
In each of the two states, different groups of FMU variables can be assigned a value. The groups are defined by FMI based on predicates on the characteristics of the variables (\textit{causality}, \textit{initial} and \textit{variability}). These predicates have been extracted and used in the implementation. 
An example of a group is the \textit{INI}-group that consists of FMU variables with $initial = exact\, \lor initial = \,approx $  and $variability \neq constant$. All variables of this group should be set in the \textit{Instantiated}-phase of a co-simulation, and since they have no dependencies/connections to other FMU variables, the order is insignificant. However, all variables of each FMU are group together to perform the fewest possible operations in the initialization.

@@

\subsection{Method to calculate the initialization order}
This section describes the approach to calculate the initialization order of the interconnected FMU variables are based on the the method proposed in Gomes et al. \cite{Gomes2019b, BromanCompositionCo-Simulation}, but the approach used here has been extended with the ability to handle algebraic loops. The approach is formulated in pseudo-code in algorithm \ref{alg:initialization}.
\begin{figure}[htb]
  \centering
    \begin{algorithm}[H]
    \caption{Initialization strategy for Interconnected variables}
    \label{alg:initialization}
      \begin{algorithmic}[1]
        \State $InitializationGraph \gets createGraph(connections)$
        \State $SCCS \gets Tarjan(InitializationGraph)$
        \State $OptimizeInitializationOrder(SCCS)$
        \ForEach {$SCC \in SCCS$}
            \If {$isAlgebraicLoop(SCC)$}
                \State $applyFixedPointIteration(SCC)$;
            \Else
                \State $initializeVariable(SCC)$;
            \EndIf
        \EndFor
        \end{algorithmic}
    \end{algorithm}
\end{figure}

@@

>>>>

\section{Calculation of an Initialization Order}\label{sc:initilization}
The FMI specification defines certain information about the initialization order described through different states of a co-simulation. The initialization phase covers the two states (in chronological order):
\begin{itemize}
    \item \textit{Instantiated}
    \item \textit{Initialization Mode}
\end{itemize}
In each of the two states, different groups of FMU variables can be assigned a value. The groups are defined by FMI based on rules on the characteristics of the variables (\textit{causality}, \textit{initial} and \textit{variability}). These rules have been extracted as predicates and used in the implementation. 
An example of a group from the FMI specification is the \textit{INI}-group that consists of FMU variables with $initial = exact\, \lor initial = \,approx $  and $variability \neq constant$. All variables of this group should be set in the \textit{Instantiated}-phase of a co-simulation, and since they have no dependencies/connections to other FMU variables, the order is insignificant. However, all variables of each FMU are grouped together to perform the fewest possible operations in the initialization.

@@

\subsection{Method to calculate the initialization order}
This section describes the approach to calculate the initialization order of the interconnected FMU variables are based on the the method proposed in Gomes et al. \cite{Gomes2019b, BromanCompositionCo-Simulation}, but the approach used here has been extended with the ability to handle algebraic loops.

@@

The entire approach of the initialization of interconnected variables is formulated in pseudo-code in algorithm \ref{alg:initialization}.
\begin{figure}[htb]
  \centering
    \begin{algorithm}[H]
    \caption{Initialization strategy for Interconnected variables}
    \label{alg:initialization}
      \begin{algorithmic}[1]
        \State $InitializationGraph \gets createGraph(connections)$
        \State $SCCS \gets Tarjan(InitializationGraph)$
        \State $OptimizeInitializationOrder(SCCS)$
        \ForEach {$SCC \in SCCS$}
            \If {$isAlgebraicLoop(SCC)$}
                \State $applyFixedPointIteration(SCC)$;
            \Else
                \State $initializeVariable(SCC)$;
            \EndIf
        \EndFor
        \end{algorithmic}
    \end{algorithm}
\end{figure}

### 2020-06-29 introduction.tex
Added some based on the feedback from Christian

A typical co-simulation consists of three phases: Initialization, simulation, and teardown. This work concentrates on the first of these phases - the initialization phase. The FMI specification specifies criteria for how a single FMU shall be initialized. However, the FMI standard is not concerned with how a connected system of multiple FMUs should be correctly initialized as a whole. 
The way a system of multiple FMUs should be initialized and interacted with depends on each FMUs implementation and interconnections to other FMUs \cite{gomes_lucio_vangheluwe_2019}, since these place precedence constraints between the FMU variables. The precedence constraints between the variables can introduce algebraic loops, which places particular requirements on the initialization order since the algebraic loops require a particular initialization strategy to be applied to calculate the initial values of the variables in the algebraic loop\cite{Bastian2011a}. The order of initialization is vital for the correctness of the co-simulation \cite{Thule2018} because it ensures a variable is never read before it is set like the classical \textit{readers–writers problem} and to make sure that all initial values have stabilized at the start of the simulation.

\noindent Algebraic loops occur whenever there is an interconnected FMU variable that indirectly depends on itself. Algebraic loops introduce specific challenges in the co-simulation, and studies have shown that neglecting a loop can lead to a prohibitively high error in the co-simulation \cite{Kubler2000}. Other approaches for the generation of co-simulation algorithms have avoided co-simulation scenarios containing algebraic loops since their presence reduces the chance of obtaining a valid co-simulation result\cite{Amalio2016CheckingCo-simulation, BromanCompositionCo-Simulation}. This choice is motivated by the fact that not all co-simulation scenarios containing algebraic loops are valid since some algebraic loops have a divergent nature. However, experience has shown that valid co-simulation scenarios containing algebraic loops exist, and a well-established co-simulation framework should be able to handle these scenarios.

>>>>

A typical co-simulation consists of three phases: Initialization, simulation, and teardown. This work concentrates on the first of these phases - the initialization phase. The FMI specification specifies criteria for how a single FMU shall be initialized. However, the FMI standard is not concerned with how a connected system of multiple FMUs should be correctly initialized as a whole. 
The way a system of multiple FMUs should be initialized and interacted with depends on each FMUs implementation and interconnections to other FMUs \cite{gomes_lucio_vangheluwe_2019}, since these place precedence constraints between the FMU variables. The precedence constraints between the variables can introduce algebraic loops, which places particular requirements on the initialization order since the algebraic loops require a particular initialization strategy to be applied to calculate the initial values of the variables in the algebraic loop\cite{Bastian2011a}. An algebraic loops occur whenever there is an interconnected FMU variable that indirectly depends on itself. Algebraic loops introduce specific challenges in the co-simulation, and studies have shown that neglecting a loop can lead to a prohibitively high error in the co-simulation \cite{Kubler2000}. The order of initialization is vital for ensuring the correctness of the co-simulation \cite{Thule2018} because it ensures a variable is never read before it is set like the classical \textit{readers–writers problem} and to make sure that all initial values have stabilized at the start of the simulation.

\noindent Other approaches for the generation of co-simulation algorithms have avoided co-simulation scenarios containing algebraic loops since their presence reduces the chance of obtaining a valid co-simulation result\cite{Amalio2016CheckingCo-simulation, BromanCompositionCo-Simulation}. This choice is motivated by the fact that not all co-simulation scenarios containing algebraic loops are valid since some algebraic loops have a divergent nature. However, experience has shown that valid co-simulation scenarios containing algebraic loops exist like \todo{Claudio can you help me come up with an example? Maybe some kind of feedback control}, and a well-established co-simulation framework should be able to handle these scenarios.
